---
title: "Lesson 1"
author: "Cong Wang & Renan Serenini"
date: "2022-08-25"
draft: false
excerpt: 
weight: 1
links:
- icon: github
  icon_pack: fab
  name: code
  url: https://github.com/CongWang141/statistics-with-r.git
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### author
Cong Wang & Renan Serenini

## Set language
Set system language to English. Useful to search for error solutions
```{r}
Sys.setenv(LANG = "en")
```

To simplify the work with R,  you can set working directory. Shortcut: Ctrl + Shift + H
```{r}
setwd("C:/...")
```

## Basic skills for R
### Basic operations
With R you can do some simple calculations such as:
```{r}
3+4
sqrt(25)
6^2
3**3
```

Two instructions can be placed on the same row separeted by ";"
```{r}
4 * 2; 5 ^ 2
```

### Objects
Usual assignment operator in R. Shortcut: Alt + "-",  by doing this we simply assignment the value of **a** to **3**.
```{r}
a <- 3
```

### Vectors
Vector of numbers
```{r}
x <- c(10.4, 5.6, 3.1, 6.4, 21.7)
v1 <- c(1:10)
```

Abstract some elements from vector
```{r}
x[1]
```

Operations with vectors
```{r}
1/x
```

Create a new vector,  using an existing operations
```{r}
y <- c(x, 0, x)
```

Vectors arithemetic 
```{r}
v <- 2*x + 3*x
v
```

Mean of vector x
```{r}
sum(x) / length(x)
```

Use a function to get the mean of vector x
```{r}
mean(x)
```

We can calculate the variance of the vector **x**, remind the variance formula $S^2=\frac{\sum{(x_i-\bar{x})}}{n-1}$ 
```{r}
sum((x-mean(x))^2)/(length(x)-1)
```

Or we can use a function
```{r}
var(x)
```

We can also creat a vectorof strings.
```{r}
s <- c('age', 'height', 'weight')
```

Create a vector by using ``sequence``
```{r}
s1 <- seq(length=10, from=2,  by=1)
s2 <- seq(3, 13, 1)
```

Create a vector by using ``repeat``
```{r}
s3 <- rep(3, times=5)
s4 <- rep(x, 5)
```

Use ``replicate`` to create a vector
```{r}
x <- 3
replicate(n=10, expr = x+1)
```

Use ``replicate`` to creat a matrix
```{r}
v <- c(2, 3, 4)
replicate(n=10, expr = v)
```

### Label a vector
use function to label a vector ``names``
```{r}
country <- c(2, 4, 1, 5)
names(country) <- c("China", "Italy", "Germany", "France")
country
```

## Simple loop structure
```{r}
for (i in 1:5){
  print(i)
}
```

## Check missing value
```{r}
z <- c(1:4, NA)
```

**z** is a vector with last element as NA, we can check it out
```{r}
ind <- is.na(z)
ind
```

## Matrix and arrays
A matrix is a two-dimensional (m x n) object, use function ``matrix`` to create a matrix
```{r}
m1 <- matrix(c(1, 2, 3, 4), nrow = 2, ncol = 2)
m1
```

An array is a three-dimensional object. All elements must be of the same type, use function ``array``, 
```{r}
a1 <- array(2, dim = c(1, 2, 3))
a1
```

```{r}
vector <- c(1:12)
multi_array <- array(vector, dim = c(2, 3, 2))
multi_array
```

Matrix operations, element by element product
```{r}
m2 <- matrix(c(2, 3, 4, 5), 2, 2)
m1*m2
```

Matrix multiplication
```{r}
m1%*%m2
```

Transpose
```{r}
t(m1)
```

Get the inverce of matrix
```{r}
solve(m1)
```

Solve equation system
\begin{equation}
\begin{cases}
5x + 6y = 12\\
2x + 8y = 10
\end{cases}
\end{equation}
```{r}
a <- matrix(c(5, 6, 2, 8), 2, 2); a
y <- c(12, 10); y
solve(a, y)
```

## Dataframe
Dataframe is similar to a matrix, but each column is independent. Elements must be of the same type just in each column.
```{r}
students <-   data.frame (index = 1:10,
              height=c(1.2, 1.3, 1.1, 1.35, 1.21,1.32, 1.33, 1.22, 1.40, 1.32),
              gender=c("male", "female ", "female", "male", "female","female", "male", "female","male","male"),
              bus_transport =c(TRUE, TRUE, FALSE, FALSE, FALSE, TRUE, TRUE,FALSE, TRUE, TRUE))
students
```

check the names of each column
```{r}
names(students)
```

Create another variable ``weight`` to the Dataframe
```{r}
students$weight <- c(40,38,28,50,34, 43, 38, 33 ,45, 37)
```

plot students' height and weight
```{r}
plot(students$height, students$weight)
```

Several arguments can be passed to improve the plot
```{r}
plot(students$height, students$weight, pch=15, col='blue')
```

## Load and save data
install a package we need to use, by using ``install.packages("openxlsx")``
loading the package
```{r}
library(openxlsx)
```

Save the dataframe we used before ``students`` in xlsx format
```{r}
write.xlsx(students, file = "students.xlsx")
```

Save the dataframe we used before ``students`` in csv format
```{r}
write.csv(students, file = "students.csv")
```

Clear the workspace by using the code below. Shortcut: Ctrl + L
```{r}
rm(list = ls())
```

Load data from an online source
```{r}
accidents=read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/airline-safety/airline-safety.csv")
```

load from the device, "Import Dataset" button is highly recommended!
students <- read.xlsx("students.xlsx")

## Data manipulation
filter the data, create a sub Dataset with ``fatal_accidents_85_99==0`` name it as ``no_accident``,  create another sub Dataset with ``fatalities_00_14>100`` and name it as ``fatalities``, you can check those new datasets by yourself
```{r}
no_accident <- subset(accidents, accidents$fatal_accidents_85_99==0)
fatalities <- subset(accidents, fatalities_00_14>100)
```

## Exercise 1
The first exercise we use daily air quality measurements in NY, May to September 1973
```{r}
data("airquality")
```

1. you are required to drop observations with one or more missing values and create a new dataset 
2. plot a line graph with temperature over time

## Exercise 2
Here is a random sample of male and female named **x1**, and a random sample from normal distribution with mean 1.7 and standared deviation 0.1 named **x2**, **x3** is depends on **x2** plus a random error with mean=1, sd=5
```{r}
x1 <- sample(c('male', 'female'),1000, replace = TRUE)
x2 <- rnorm(1000, 1.7, 0.1)
x3 <- x2*45 + rnorm(1000, 1, 5)
```
1. you are required to create a dataset based on those vectors above
2. Name columns as x1=Gender, x2=Height, x3=Weight
3. Create  a new column BMI which denotes Body mass Index with the following formula: bmi= weight/height^2
4. Suppose obese condition applies to male if BMI>28 or female IF BMI>30, filter observations with obese condition and make a scatterplot (Weight ~ Height)