---
title: "Lesson 6"
author: "Cong Wang & Renan Serenini"
date: "2022-08-25"
draft: false
excerpt: 
weight: 1
links:
- icon: github
  icon_pack: fab
  name: code
  url: https://github.com/CongWang141/statistics-with-r.git
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### author
Cong Wang & Renan Serenini

## Solutions to lesson 5

1. Load the dataset 'mtcars' (check the help for information about it)

2. Check the descriptive statistics of the variables and the correlation

```{r, message=FALSE}
data(mtcars)
?mtcars
```

3. Estimate a multiple linear model to identify the variables that impact the fuel consumption of a car

```{r}
attach(mtcars)
model1 <- lm(mpg~cyl+wt)
summary(model1)
plot(model1$residuals)
```

4. Test the assumption of heteroskedasticity

```{r, message=FALSE}
library(lmtest)
bptest(model1) #NUll: The error variances are all equal
```

5. Interpret the results of the model

6. From the same model, include another variable and repeat the tests and interpretations

```{r}
#New variable
model2 <- lm(mpg~cyl+wt+gear)
summary(model2)
```

7. What was the effect of including this new variable? Would you keep it in the model?

8. Create a dummy variable and include it in your model, commenting the result.

```{r, message=FALSE}
#Create a dummy variable
mtcars$carburetors <- ifelse(carb>2,1,0)
attach(mtcars)
model3 <- lm(mpg~cyl+wt+carburetors)
summary(model3)
```

9. Generate random data for a new car and predict the values of it's fuel consumption based on your model

```{r}
#Predict a random car
Panda <- cbind('cyl'=numeric(1), 'wt'=numeric(1))
Panda[1] <-   sample(mtcars$cyl, size=1)+rnorm(1) #Sample + random error term
Panda[2] <-   sample(mtcars$wt, size=1)+rnorm(1)  
Panda <- data.frame(Panda) #Transform into data frame to use function predict.lm
predict.lm(model1, Panda)
```

check the function ``predict`` and ``predict.lm`` if you are not familiar with them
```{r}
?predict
?predict.lm
```

## Hypothesis Tests in Multiple Regression 

```{r, message=FALSE}
library(AER)
library(stargazer)
data(CASchools)
```

```{r}
#Our new variables
CASchools$STR <- CASchools$students/CASchools$teachers #STR of classes       
CASchools$score <- (CASchools$read + CASchools$math)/2 #Score

model <- lm(score ~ STR + english, data = CASchools)
coeftest(model, vcov. = vcovHC, type = "HC1")
```

```{r}
# compute two-sided p-value MANUALLY
t_b1=coeftest(model, vcov. = vcovHC, type = "HC1")[2, 3];t_b1 #t_statistic
p_b1=pt(abs(t_b1),df = model$df.residual); p_b1 #Distribution functions
p_value_b1=2 * (1 - p_b1); p_value_b1
```

## Confidence Intervals in Multiple Regression 

warning: confint does not use robust standard errors

```{r}
model <- lm(score ~ STR + english, data = CASchools)
confint(model, level = 0.9)
```

```{r}
# compute robust standard errors
rob_se <- diag(vcovHC(model, type = "HC1"))^0.5
rob_se
```

```{r}
# compute robust 95% confidence intervals
rbind("lower" = coef(model) - qnorm(0.975) * rob_se,
      "upper" = coef(model) + qnorm(0.975) * rob_se)
```

```{r}
# compute robust 90% confidence intervals
rbind("lower" = coef(model) - qnorm(0.95) * rob_se,
      "upper" = coef(model) + qnorm(0.95) * rob_se)
```

Let's include another variable
```{r}
# scale expenditure to thousands of dollars
CASchools$expenditure <- CASchools$expenditure/1000
```

```{r}
# estimate the model
model <- lm(score ~ STR + english + expenditure, data = CASchools)
summary(model)
coeftest(model, vcov. = vcovHC, type = "HC1")
```

```{r}
# Multicolinearity issue
cor(CASchools$STR, CASchools$expenditure)
```

## F-Statistic 

```{r}
# estimate the multiple regression model
model <- lm(score ~ STR + english + expenditure, data = CASchools)
```

Execute the function on the model object and provide both linear restrictions 

```{r}
linearHypothesis(model, c("STR=0", "expenditure=0"))
#NULL: B1=B2=0

# heteroskedasticity-robust F-test
linearHypothesis(model, c("STR=0", "expenditure=0"), white.adjust = "hc1")
```

```{r}
#Overall F-test
linearHypothesis(model, c("STR=0", "english=0", "expenditure=0"), white.adjust = "hc1")
summary(model)
summary(model)$fstatistic
ls(summary(model))
```

## Confidence Sets for Multiple Coefficients 

draw the 95% confidence set for coefficients on STR and expenditure
```{r}
confidenceEllipse(model, 
                  fill = T,
                  lwd = 0,
                  which.coef = c("STR", "expenditure"),
                  main = "95% Confidence Set")
```

draw the **robust** 95% confidence set for coefficients on STR and expenditure, and the 95% confidence set for coefficients on STR and expenditure
```{r}
confidenceEllipse(model, 
                  fill = T,
                  lwd = 0,
                  which.coef = c("STR", "expenditure"),
                  main = "95% Confidence Sets",
                  vcov. = vcovHC(model, type = "HC1"),
                  col = "red")

confidenceEllipse(model, 
                  fill = T,
                  lwd = 0,
                  which.coef = c("STR", "expenditure"),
                  add = T) #Argument add=T (TRUE)
```

