---
title: "Global Carbon Emission Analysis"
author: "Cong Wang"
date: "2022-07-30"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float: true
excerpt: This airticle analyize the globle carobon emission. The data for this analysis is from **World Bank Open Data**, collected since 1990. To break down the global carbon emission and decompose the trend.
tags: 
- Carbon Emission

categories: 
- Climate Change

featured: true
layout: single-sidebar
---
<style type="text/css">
  h1,h2,h3{
  font-size: 16pt;
  color: steelblue
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Sourse
The data for this analysis is from **World Bank Open Data**, link: https://data.worldbank.org/

  To download the [CSV.file](https://api.worldbank.org/v2/en/indicator/EN.ATM.CO2E.KT?downloadformat=csv), you will get:
  
1. each country's carbon emission data
2. country metadata
3. indicator metadata

# Data exploration 
```{r, echo=FALSE}
setwd("D:/ComputationalTool/RStudy/GlobalClimateChange")
```

Import each country's carbon emission data.
```{r}
d <- read.csv("API_EN.ATM.CO2E.KT_DS2_en_csv_v2_4354173.csv", skip = 3)
```

Carbon emission data is collected since 1990, the year before that shows NA, we need to remove the year without data and remove variables we do not need.  
```{r, message=FALSE}
library(tidyverse)
```

```{r}
not_all_na <- function(x) any(!is.na(x))
d_comp <- select(d, where(not_all_na),  
                            -"Indicator.Name", -"Indicator.Code")
```

## Co2 emission for the whole world.{.tabset}
The world total carbon emission is contained in the data, with country name *World*.

Pick out the data for **World**, convert it into panel data.
```{r}
world <- d_comp[d_comp$Country.Name=="World", ]
world_p <- pivot_longer(world, names_to="year", values_to = "co2", cols = starts_with("X"))
```

Remove the X in front of each year, and declare it is number.
```{r}
world_p$year <- as.numeric(substr(world_p$year, 2, 5)) 
```


### Use ggplot2.{-}
```{r}
library(ggplot2)
p1 <- ggplot(data = world_p, aes(x=year, y=co2/1000000))
p1 + geom_line(color="steelblue") + 
  scale_x_continuous(breaks=seq(1988,2021,2))+
  scale_y_continuous(breaks=seq(17, 34, 2)) + ylab("Co2") + 
  ggtitle("in million")
```

### YOY Change in percentage.{-}
Calculate the year over year change of carbon emission.
```{r}
world_p$change <- (world_p$co2-dplyr::lag(world_p$co2))/dplyr::lag(world_p$co2)*100 
```

```{r, warning = FALSE}
p1.1 <- ggplot(data = world_p, aes(x=year, y=change))
p1.1 + geom_line(color="steelblue") + 
  scale_x_continuous(breaks=seq(1988,2021,2))+
ylab("YOY Change of Co2 Emission") + 
  ggtitle("In Percentage")
```

## Co2 emission in different area{.tabset}
The dataset also includes carbon emission data for different geographic areas.

Get total carbon emission for each row through all time
```{r}
d_comp$co2_total <- rowSums(d_comp[ ,3:32])
d_comp <- arrange(d_comp, desc(co2_total)) 
```

Except for the world total carbon emission, different geographic area's carbon emission is also collected in the data. Pick geographic area.
```{r}
co2_area <- d_comp[d_comp$Country.Name %in% c("East Asia & Pacific", "North America", "Euro area", "Middle East & North Africa", "South Asia", "Latin America & Caribbean", "Sub-Saharan Africa"), ]

```

Convert dataset to panel data.
```{r}
co2_area_p <- pivot_longer(co2_area, names_to="year", values_to = "co2", cols = starts_with("X"))
co2_area_p$year <- as.numeric(substr(co2_area_p$year, 2, 5)) 
co2_area_p <- arrange(co2_area_p, desc(co2_total))

```

### Line Plot.{-}
```{r}
p2 <- ggplot(data = co2_area_p, aes(x=year, y=co2/1000000, color=Country.Name))
p2 + geom_line() + 
  scale_x_continuous(breaks=seq(1988,2021,4))+
  scale_y_continuous(breaks=seq(0, 16, 2)) + ylab("Co2") + 
  ggtitle("in million") +
  theme(legend.position="bottom", legend.title = element_blank())
```


### Bar Plot.{-}
```{r}
p3 <- ggplot(data = co2_area_p, aes(x=reorder(Country.Name, -co2_total), y=co2_total/1000000))
p3 + geom_bar(stat="identity", fill="steelblue") + 
  ggtitle("in million") +  ylab("Total Co2 Emission") + xlab("Country") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) 
```

### YOY Change{-}
Group the data *co2_area_p* by country name, in each group calculate the year over year carbon emission change in percentage.
```{r, message=FALSE}
co2_area_p %>%
  group_by(Country.Name) %>%
  summarise((co2-dplyr::lag(co2))/dplyr::lag(co2)*100) -> co2_area_p$change
```
By using "group_by" and "summarise", R will generate a dataset under variable "change", we need to handle it. 
```{r}
co2_area_p$change <- co2_area_p$change$`(co2 - dplyr::lag(co2))/dplyr::lag(co2) * 100`
```


Plot the data
```{r, warning = FALSE}
p2.1 <- ggplot(data = co2_area_p, aes(x=year, y=change, color=Country.Name))
p2.1 + geom_line() + 
  scale_x_continuous(breaks=seq(1988,2021,4))+
ylab("YOY Co2 Emission Change") + 
  ggtitle("In Percentage") +
  theme(legend.position="bottom", legend.title = element_blank())
```



## Co2 emission for top 10 countries{.tabset}
Get carbon emission for top 10 countries through all time.
```{r}
country_data <- d_comp[c(15, 17, 30, 32, 36, 43:266), ]
country_10 <- country_data[c(1:8, 10:11), ]
```

Convert it to panel data
```{r}
country_10_pan <- pivot_longer(country_10, names_to="year", values_to = "co2", cols = starts_with("X"))
country_10_pan$year <- as.numeric(substr(country_10_pan$year, 2, 5)) 

```

### Line Plot for top 10 countries.{-}
```{r}
p4 <- ggplot(data = country_10_pan, aes(x=year, y=co2/1000000, color=Country.Name))
p4 + geom_line() + 
  scale_x_continuous(breaks=seq(1988,2021,4))+
  scale_y_continuous(breaks=seq(0, 16, 2)) + ylab("Co2") + 
  ggtitle("in million") +
  theme(legend.position="bottom", legend.title = element_blank())
```

### Bar Plot for top 10 countries.{-}
```{r}
p5 <- ggplot(data = country_10_pan, aes(x=reorder(Country.Name, -co2_total), y=co2_total/1000000))
p5 + geom_bar(stat="identity", fill="steelblue") + 
  ggtitle("in million") +  ylab("Total Co2 Emission") + xlab("Country") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) 
```

### YoY Change for top 10 countries.{-}

```{r, message=FALSE}
country_10_pan %>%
  group_by(Country.Name) %>%
  summarise((co2-dplyr::lag(co2))/dplyr::lag(co2)*100) -> country_10_pan$change
```

```{r}
country_10_pan$change <- country_10_pan$change$`(co2 - dplyr::lag(co2))/dplyr::lag(co2) * 100`
```

Plot the data
```{r, warning = FALSE}
p2.1 <- ggplot(data = country_10_pan, aes(x=year, y=change, color=Country.Name))
p2.1 + geom_line() + 
  scale_x_continuous(breaks=seq(1988,2021,4))+
ylab("YOY Co2 Emission Change") + 
  ggtitle("In Percentage") +
  theme(legend.position="bottom", legend.title = element_blank())
```

### YoY Change for top 5 countries.{-}
```{r, warning = FALSE}
p2.2 <- ggplot(data = country_10_pan[1:150, ], aes(x=year, y=change, color=Country.Name))
p2.2 + geom_line() + 
  scale_x_continuous(breaks=seq(1988,2021,4))+
ylab("YOY Co2 Emission Change") + 
  ggtitle("In Percentage") +
  theme(legend.position="bottom", legend.title = element_blank())
```













